\documentclass[12pt,a4paper]{article}

% ─── Packages ───────────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{tcolorbox}
\usepackage{parskip}
\usepackage{fontawesome5}
\usepackage{lmodern}

% ─── Colors ─────────────────────────────────────────────────────────────────
\definecolor{primaryblue}{RGB}{37,99,235}
\definecolor{darkblue}{RGB}{30,64,175}
\definecolor{codebg}{RGB}{248,250,252}
\definecolor{codeframe}{RGB}{226,232,240}
\definecolor{terminalgreen}{RGB}{16,185,129}
\definecolor{darkorange}{RGB}{234,88,12}
\definecolor{rose}{RGB}{244,63,94}
\definecolor{textgray}{RGB}{71,85,105}

% ─── Code Listings ──────────────────────────────────────────────────────────
\lstset{
  backgroundcolor=\color{codebg},
  frame=single,
  framerule=0.5pt,
  rulecolor=\color{codeframe},
  basicstyle=\ttfamily\small,
  keywordstyle=\color{primaryblue}\bfseries,
  commentstyle=\color{textgray}\itshape,
  stringstyle=\color{darkorange},
  numbers=left,
  numberstyle=\tiny\color{textgray},
  stepnumber=1,
  tabsize=4,
  showspaces=false,
  showstringspaces=false,
  breaklines=true,
  captionpos=b
}

\lstdefinelanguage{bash}{
  keywords={python,pip,git,cd,npm,npx},
  morecomment=[l]{\#},
  morestring=[b]",
}

% ─── Section styling ────────────────────────────────────────────────────────
\titleformat{\section}{\Large\bfseries\color{primaryblue}}{}{0em}{}[\color{codeframe}\titlerule]
\titleformat{\subsection}{\large\bfseries\color{darkblue}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{textgray}}{}{0em}{}

% ─── Callout boxes ──────────────────────────────────────────────────────────
\tcbuselibrary{skins,breakable}
\newtcolorbox{notebox}{colback=blue!5,colframe=primaryblue,fonttitle=\bfseries,title=Note}
\newtcolorbox{warnbox}{colback=orange!5,colframe=darkorange,fonttitle=\bfseries,title=Important}
\newtcolorbox{termbox}{colback=black!95,colframe=black,colupper=terminalgreen,fontupper=\ttfamily\small,title=Terminal}

% ─── Header / Footer ────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\rhead{\color{textgray}\small TaskFlow — Technical Documentation}
\lhead{\color{primaryblue}\small\textbf{TF}}
\rfoot{\color{textgray}\small Page \thepage}

% ─── Hyperlink setup ────────────────────────────────────────────────────────
\hypersetup{
    colorlinks=true,
    linkcolor=primaryblue,
    urlcolor=primaryblue,
    pdftitle={TaskFlow Technical Documentation},
    pdfauthor={TaskFlow Team},
}

% ════════════════════════════════════════════════════════════════════════════
\begin{document}
% ════════════════════════════════════════════════════════════════════════════

\begin{titlepage}
    \centering
    \vspace*{3cm}
    {\Huge\bfseries\color{primaryblue} TaskFlow}\\[0.5em]
    {\Large\color{textgray} Complete Technical Documentation}\\[2em]
    \noindent\rule{\linewidth}{0.5pt}\\[2em]
    {\large A full-stack Team Task Manager built with\\[0.3em]
    \textbf{Django REST Framework} + \textbf{React}}\\[3em]
    {\color{textgray} Covers: Architecture · Backend · Frontend · API · Deployment}\\[1em]
    {\small\color{textgray} Version 1.0 · March 2026}
    \vfill
\end{titlepage}

\tableofcontents
\newpage


% ════════════════════════════════════════════════════════════════════════════
\section{Project Overview}
% ════════════════════════════════════════════════════════════════════════════

TaskFlow is a real-time, collaborative task management application. It allows teams to create workspaces, assign tasks to multiple members, and track progress — all from a clean, modern interface.

\subsection{High-Level Architecture}

The project is split into two completely separate applications that communicate over HTTP:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Layer} & \textbf{Technology} & \textbf{Responsibility} \\
\midrule
Frontend & React (Vite) & User Interface \\
Backend & Django (Python) & Business Logic + Database \\
Database & PostgreSQL (Neon.tech) & Data Storage \\
\bottomrule
\end{tabular}
\end{center}

\begin{notebox}
The Frontend and Backend are \textbf{completely independent}. The Frontend does NOT touch Python or Django at all. They talk to each other only through \textbf{API calls} (HTTP requests sending JSON data).
\end{notebox}

\subsection{Live URLs}
\begin{itemize}
    \item \textbf{Frontend (Vercel):} \url{https://task-flow-theta-six.vercel.app}
    \item \textbf{Backend API (Render):} \url{https://task-manager-backend-zfld.onrender.com}
    \item \textbf{Database (Neon.tech):} PostgreSQL cloud database
\end{itemize}

\subsection{Project Folder Structure}
\begin{lstlisting}
TaskFlow/                  <- root repo
├── backend/               <- Django project
│   ├── core/              <- settings, URLs, WSGI
│   │   ├── settings.py    <- all configuration
│   │   ├── urls.py        <- URL routing (top-level)
│   │   └── wsgi.py        <- production server entry
│   ├── api/               <- our main app
│   │   ├── models.py      <- database tables
│   │   ├── serializers.py <- data validation & formatting
│   │   ├── views.py       <- request handlers (API logic)
│   │   ├── urls.py        <- API URL routing
│   │   ├── permissions.py <- access control rules
│   │   └── exceptions.py  <- custom error handling
│   ├── requirements.txt   <- Python libraries
│   └── .env               <- secret environment variables
│
└── frontend/              <- React project
    ├── src/
    │   ├── components/    <- reusable UI pieces
    │   ├── pages/         <- full pages
    │   ├── store/         <- global state (Zustand)
    │   ├── lib/           <- API client (Axios)
    │   └── index.css      <- global styles
    ├── public/            <- static assets (logo, etc.)
    └── index.html         <- entry HTML file
\end{lstlisting}


% ════════════════════════════════════════════════════════════════════════════
\section{Technology Stack Explained}
% ════════════════════════════════════════════════════════════════════════════

\subsection{Backend Technologies}

\subsubsection{Python}
Python is the programming language the backend is written in. It reads almost like plain English, which makes it great for beginners. We use Python 3.11+.

\subsubsection{Django}
Django is a high-level \textbf{web framework} for Python. Think of it as a toolkit that handles:
\begin{itemize}
    \item Connecting to the database
    \item Handling user authentication and sessions
    \item URL routing (deciding which code runs when someone visits a URL)
    \item Admin panel (a built-in dashboard for managing data)
\end{itemize}

\subsubsection{Django REST Framework (DRF)}
DRF is a plugin on top of Django. It converts Django into an \textbf{API server} — instead of returning HTML pages, it returns \textbf{JSON data}. The frontend then reads this JSON.

\subsubsection{PostgreSQL}
A powerful, open-source database. We store all user data, teams, tasks, and notifications here. Django talks to it using \textbf{models} (Python classes that represent tables).

\subsubsection{Gunicorn}
A \textbf{production web server} for Python apps. In development we use Django's built-in server, but in production on Render we use Gunicorn because it's faster and more stable.

\subsubsection{Whitenoise}
Serves static files (CSS, images) directly from Django without needing a separate file server.

\subsubsection{Key Python Libraries}
\begin{center}
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Library} & \textbf{Version} & \textbf{Purpose} \\
\midrule
django & 5.0.6 & Core web framework \\
djangorestframework & 3.15.2 & JSON API layer \\
django-cors-headers & 4.4.0 & Allow cross-domain requests from the frontend \\
psycopg2-binary & 2.9.9 & PostgreSQL database driver for Python \\
python-dotenv & 1.0.1 & Load secrets from a .env file \\
whitenoise & 6.7.0 & Serve static files \\
gunicorn & 22.0.0 & Production WSGI server \\
dj-database-url & 2.1.0 & Parse database URL from environment variable \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Frontend Technologies}

\subsubsection{React (with Vite)}
React is a JavaScript library for building user interfaces. \textbf{Vite} is the build tool that bundles all the JavaScript/CSS files and runs a fast development server.

\subsubsection{Tailwind CSS}
A CSS framework where you style elements using short utility class names directly in your HTML/JSX (e.g., \texttt{bg-blue-50 text-slate-900 rounded-xl}).

\subsubsection{TanStack Query (React Query)}
Manages server state — handles fetching data from the API, caching it, and automatically refetching when needed. Replaces manual \texttt{useEffect + fetch} code.

\subsubsection{Zustand}
A minimal global state manager. Used to store the logged-in user's information so any component can access it without passing it as a prop.

\subsubsection{Axios}
An HTTP client library. Used to make API calls to the Django backend. We have a custom instance \texttt{api.js} that automatically adds auth headers.

\subsubsection{React Hook Form + Zod}
Handles form validation. Zod defines the schema (rules), and React Hook Form manages the form state efficiently.

\subsubsection{Lucide React}
Icon library — provides all the icons used in the UI (bell, trash, edit, etc.).


% ════════════════════════════════════════════════════════════════════════════
\section{Backend — Deep Dive}
% ════════════════════════════════════════════════════════════════════════════

\subsection{Django Concepts Explained Simply}

\subsubsection{What is a Model?}
A \textbf{Model} is a Python class that represents a \textbf{database table}. Each attribute of the class is a column in that table. When you define a model, Django can automatically create the database table for you using \textbf{migrations}.

\begin{lstlisting}[language=Python, caption=Example: a simple model]
class Task(models.Model):
    title = models.CharField(max_length=200)   # text column
    status = models.CharField(...)              # another text column
    due_date = models.DateField(null=True)      # date column, can be empty
\end{lstlisting}

\subsubsection{What is a Migration?}
A migration is like a \textbf{changelog} for your database. When you change a model (add/remove a field), you run two commands:
\begin{termbox}
# Step 1: Generate the migration file (detects your changes)
python manage.py makemigrations

# Step 2: Apply it to the database
python manage.py migrate
\end{termbox}
Django tracks which migrations have been applied and only runs new ones.

\subsubsection{What is a Serializer?}
A serializer converts \textbf{Python objects} (Django model instances) into \textbf{JSON} (for sending to the frontend), and vice versa. It also validates incoming data.

\begin{lstlisting}[language=Python, caption=Serializer flow]
# Incoming JSON from frontend:
{ "title": "Fix bug", "priority": "high", "team": 1 }
         |
         v
# Serializer validates it, converts to Python dict
         |
         v
# .save() creates/updates a Task object in the database
\end{lstlisting}

\subsubsection{What is a ViewSet?}
A ViewSet is a class that handles multiple related HTTP methods in one place. For example, \texttt{TaskViewSet} handles:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Method} & \textbf{URL} & \textbf{Action} \\
\midrule
GET & /api/tasks/ & List all tasks \\
POST & /api/tasks/ & Create a new task \\
GET & /api/tasks/5/ & Get task with id=5 \\
PATCH & /api/tasks/5/ & Update task with id=5 \\
DELETE & /api/tasks/5/ & Delete task with id=5 \\
\bottomrule
\end{tabular}
\end{center}


\subsection{Database Models}

\subsubsection{User Model}
Extends Django's built-in \texttt{AbstractUser}. We override it to use \textbf{email} as the login field instead of username.

\begin{lstlisting}[language=Python]
class User(AbstractUser):
    email = models.EmailField(unique=True)
    USERNAME_FIELD = 'email'   # login with email, not username
\end{lstlisting}

\begin{notebox}
Django stores passwords as \textbf{hashed} values using PBKDF2-SHA256. The plain password is never stored anywhere.
\end{notebox}

\subsubsection{Team Model}
A workspace that groups members and tasks. Every team has one creator/owner.

\begin{lstlisting}[language=Python]
class Team(models.Model):
    name = models.CharField(max_length=100)
    description = models.TextField(blank=True)
    creator = models.ForeignKey(User, on_delete=models.CASCADE)
    created_at = models.DateTimeField(auto_now_add=True)  # set once on create
    updated_at = models.DateTimeField(auto_now=True)       # updated on every save
\end{lstlisting}

\textbf{ForeignKey} means: each team has \textbf{one} creator, and that creator is a User. If the User is deleted, the team is also deleted (\texttt{CASCADE}).

\subsubsection{TeamMembership Model}
This is a \textbf{junction table} (many-to-many relationship with extra data). It connects Users to Teams and stores their role (owner/member).

\begin{lstlisting}[language=Python]
class TeamMembership(models.Model):
    team = models.ForeignKey(Team, ...)
    user = models.ForeignKey(User, ...)
    role = models.CharField(choices=[('owner','Owner'),('member','Member')])
    joined_at = models.DateTimeField(default=timezone.now)
    
    class Meta:
        unique_together = ('team', 'user')  # can't be in same team twice
\end{lstlisting}

\subsubsection{Task Model}
The core model. Tasks belong to a team, are created by one user, and can be \textbf{assigned to multiple users} (ManyToMany).

\begin{lstlisting}[language=Python]
class Task(models.Model):
    title = models.CharField(max_length=200)
    status = models.CharField(choices=['todo','in_progress','done'])
    priority = models.CharField(choices=['low','medium','high'])
    due_date = models.DateField(null=True, blank=True)
    
    team = models.ForeignKey(Team, ...)          # belongs to ONE team
    assigned_to = models.ManyToManyField(User)   # many users can be assigned
    created_by = models.ForeignKey(User, ...)    # created by ONE user
\end{lstlisting}

\textbf{ManyToManyField} automatically creates a hidden database table (e.g., \texttt{task\_assigned\_to}) with rows like \texttt{(task\_id, user\_id)}.

\subsubsection{Notification Model}
Stores per-user notifications for events like overdue tasks and task assignments.

\begin{lstlisting}[language=Python]
class Notification(models.Model):
    user = models.ForeignKey(User, ...)   # belongs to ONE user
    title = models.CharField(max_length=200)
    message = models.TextField()
    type = models.CharField(choices=['overdue_task','task_assignment', ...])
    is_read = models.BooleanField(default=False)
\end{lstlisting}


\subsection{Authentication — How Login Works}

We use \textbf{Session-based authentication}, which works like this:

\begin{enumerate}
    \item User sends email + password to \texttt{POST /api/auth/login/}
    \item Django validates credentials using \texttt{authenticate()}
    \item \texttt{login(request, user)} creates a \textbf{session} in the database and sends a \textbf{session cookie} to the browser
    \item On every subsequent request, the browser automatically sends the cookie
    \item Django looks up the session, finds the user, and the user is "logged in"
    \item The response also includes a \texttt{csrf\_token} for security
\end{enumerate}

\begin{notebox}
Because our frontend (Vercel) and backend (Render) are on \textbf{different domains}, cookies need special settings: \texttt{SameSite=None} and \texttt{Secure=True}. This is required for cross-origin cookie sharing.
\end{notebox}

\subsection{CSRF Protection}

\textbf{CSRF} (Cross-Site Request Forgery) is an attack where a malicious website tricks your browser into making requests to our API using your session.

Django's built-in CSRF protection works by:
\begin{enumerate}
    \item Setting a hidden \texttt{csrftoken} cookie 
    \item Requiring every POST/PATCH/DELETE request to include this token in the header
    \item Rejecting requests that don't have it
\end{enumerate}

Since cookies are unreliable across domains, we also \textbf{return the CSRF token in the login response JSON}, and the frontend stores it manually and attaches it to every request header (\texttt{X-CSRFToken}).

\subsection{Permissions and Access Control}

We have a custom permission class \texttt{IsTeamOwner} that restricts certain team actions (edit/delete team, add/remove members) to only the team owner.

\begin{lstlisting}[language=Python]
class IsTeamOwner(BasePermission):
    def has_object_permission(self, request, view, obj):
        # obj is the Team
        return TeamMembership.objects.filter(
            team=obj, user=request.user, role='owner'
        ).exists()
\end{lstlisting}

All other endpoints require \texttt{IsAuthenticated} (user must be logged in).


% ════════════════════════════════════════════════════════════════════════════
\section{API Endpoints Reference}
% ════════════════════════════════════════════════════════════════════════════

All endpoints are prefixed with \texttt{/api/}. Only auth endpoints are public; all others require a valid session cookie.

\subsection{Authentication Endpoints}
\begin{center}
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Description} \\
\midrule
POST & /api/auth/register/ & Register a new user, returns user + csrf\_token \\
POST & /api/auth/login/ & Login, returns user + csrf\_token \\
POST & /api/auth/logout/ & Destroy session \\
GET & /api/auth/me/ & Get current logged-in user \\
\bottomrule
\end{tabularx}
\end{center}

\subsubsection{Register Request Body}
\begin{lstlisting}[language=bash]
{
  "first_name": "John",
  "last_name": "Doe",
  "username": "johndoe",
  "email": "john@example.com",
  "password": "securePassword123"
}
\end{lstlisting}

\subsection{Teams Endpoints}
\begin{center}
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Description} \\
\midrule
GET & /api/teams/ & List all teams user is a member of \\
POST & /api/teams/ & Create a new team \\
GET & /api/teams/:id/ & Get a single team \\
PATCH & /api/teams/:id/ & Update team (owner only) \\
DELETE & /api/teams/:id/ & Delete team (owner only) \\
GET & /api/teams/:id/members/ & List team members \\
POST & /api/teams/:id/add\_member/ & Add member by email \\
POST & /api/teams/:id/remove\_member/ & Remove a member (with safety check) \\
\bottomrule
\end{tabularx}
\end{center}

\subsubsection{Remove Member — Safety Flow}
\begin{lstlisting}[language=bash]
# First call (force=false)
POST /api/teams/1/remove_member/
{ "user_id": 5 }

# If user has assigned tasks, returns:
{ "warning": "user_has_tasks", "detail": "User is assigned to 3 tasks." }

# Second call (force=true) — user confirmed
POST /api/teams/1/remove_member/
{ "user_id": 5, "force": true }
# This unassigns the user from all tasks, then removes from team
\end{lstlisting}

\subsection{Tasks Endpoints}
\begin{center}
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Description} \\
\midrule
GET & /api/tasks/ & List tasks (supports filters) \\
POST & /api/tasks/ & Create a task \\
GET & /api/tasks/:id/ & Get a single task \\
PATCH & /api/tasks/:id/ & Update task \\
DELETE & /api/tasks/:id/ & Delete task (creator only) \\
\bottomrule
\end{tabularx}
\end{center}

\subsubsection{Task Filters (Query Parameters)}
\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Parameter} & \textbf{Example} \\
\midrule
team & ?team=1 \\
status & ?status=in\_progress \\
priority & ?priority=high \\
search & ?search=bug \\
my\_tasks & ?my\_tasks=true \\
overdue & ?overdue=true \\
\bottomrule
\end{tabular}
\end{center}

\subsubsection{Create Task Request Body}
\begin{lstlisting}[language=bash]
{
  "title": "Fix login bug",
  "description": "Users can't login on mobile",
  "status": "todo",
  "priority": "high",
  "team": 1,
  "assigned_to_ids": [2, 3],  // list of user IDs
  "due_date": "2026-03-10"    // or null
}
\end{lstlisting}

\subsection{Dashboard \& Notifications}
\begin{center}
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Method} & \textbf{Endpoint} & \textbf{Description} \\
\midrule
GET & /api/dashboard/ & Stats: task counts, overdue tasks \\
GET & /api/notifications/ & List unread notifications \\
POST & /api/notifications/:id/read/ & Mark one as read \\
POST & /api/notifications/read\_all/ & Mark all as read \\
\bottomrule
\end{tabularx}
\end{center}


% ════════════════════════════════════════════════════════════════════════════
\section{Frontend — Deep Dive}
% ════════════════════════════════════════════════════════════════════════════

\subsection{Application Structure}

\subsubsection{Routing}
React Router manages navigation. All routes are defined in \texttt{App.jsx}:

\begin{lstlisting}[language=bash]
/login      -> LoginPage.jsx     (public)
/register   -> RegisterPage.jsx  (public)
/           -> DashboardPage.jsx (protected - must be logged in)
/tasks      -> TasksPage.jsx     (protected)
/teams      -> TeamsPage.jsx     (protected)
\end{lstlisting}

Protected routes are wrapped in a \texttt{ProtectedRoute} component that checks if the user is authenticated and redirects to \texttt{/login} if not.

\subsubsection{API Client (lib/api.js)}
All HTTP requests go through a centralized Axios instance. It automatically:
\begin{itemize}
    \item Sets the base URL to the backend
    \item Includes \texttt{withCredentials: true} (sends session cookies)
    \item Attaches \texttt{X-CSRFToken} header to every mutation request
    \item Intercepts error responses and formats them into readable messages
\end{itemize}

\subsubsection{Global State (Zustand — authStore.js)}
Stores the logged-in user's data. Any component can access it:

\begin{lstlisting}[language=bash]
const { user, setUser, clearUser } = useAuthStore()
\end{lstlisting}

On app load, it calls \texttt{GET /api/auth/me/} to restore the session if the user refreshes the page.

\subsubsection{Data Fetching (React Query)}
React Query handles all data fetching with automatic caching:

\begin{lstlisting}[language=bash]
// Fetch teams
const { data: teams } = useQuery({
    queryKey: ['teams'],           // unique cache key
    queryFn: () => api.get('/teams/').then(r => r.data)
})

// Create a team
const mutation = useMutation({
    mutationFn: (data) => api.post('/teams/', data),
    onSuccess: () => queryClient.invalidateQueries({ queryKey: ['teams'] })
})
\end{lstlisting}

\texttt{invalidateQueries} tells React Query to re-fetch the teams list after a successful mutation, keeping the UI in sync without page refresh.

\subsection{Key Pages}

\subsubsection{DashboardPage}
Fetches stats from \texttt{/api/dashboard/} and displays:
\begin{itemize}
    \item 4 stat cards: My Teams, All Tasks, My Tasks, Due Today
    \item Task status overview: Queued / Active / Done counts
    \item Overdue task reminders panel
\end{itemize}

\subsubsection{TasksPage}
\begin{itemize}
    \item Kanban-style 3-column board: To Do / In Progress / Done
    \item Filter bar (by team, status, priority, search, my tasks, overdue)
    \item Filters are stored in URL query parameters (\texttt{?status=todo}) using \texttt{useSearchParams}
    \item Only task creators see edit/delete buttons
    \item Multiple assignee avatars shown on each card
\end{itemize}

\subsubsection{TeamsPage}
\begin{itemize}
    \item Grid of team cards, each showing members
    \item Owner can edit/delete team, add/remove members
    \item Remove member flow: warns if member has assigned tasks and asks for confirmation before proceeding
    \item Removing a member unassigns them from all tasks in that team
\end{itemize}

\subsubsection{TaskModal}
A modal dialog for creating/editing tasks:
\begin{itemize}
    \item Fields: Title, Description, Team, Due Date, Status, Priority, Assign To
    \item Assignee list is fetched fresh every time (\texttt{staleTime:0})
    \item Changing team clears selected assignees (prevents cross-team assignment bug)
    \item Loading spinner while fetching members
\end{itemize}

\subsection{State Management Strategy}

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{What} & \textbf{Where stored} & \textbf{Tool} \\
\midrule
Logged-in user profile & Global state & Zustand \\
Teams list & Server cache & React Query \\
Tasks list & Server cache & React Query \\
Team members list & Server cache & React Query \\
Notifications & Server cache & React Query \\
Form data & Local component state & React Hook Form \\
Modal open/close & Local component state & useState \\
URL filters & URL query params & useSearchParams \\
\bottomrule
\end{tabular}
\end{center}

\begin{warnbox}
\textbf{Cache Invalidation Rule:} After any mutation (add/remove member, create/update/delete task), we call \texttt{queryClient.invalidateQueries()} with the relevant key. This forces React Query to re-fetch that data immediately, so the UI updates without a page refresh.
\end{warnbox}


% ════════════════════════════════════════════════════════════════════════════
\section{Development Setup (Before Production)}
% ════════════════════════════════════════════════════════════════════════════

\subsection{Prerequisites}
\begin{itemize}
    \item Python 3.11+
    \item Node.js 18+
    \item Git
    \item A PostgreSQL database (or use SQLite for local development)
\end{itemize}

\subsection{Backend Setup}

\begin{termbox}
# 1. Clone the repository
git clone https://github.com/your-org/TaskFlow.git
cd TaskFlow/backend

# 2. Create a virtual environment (isolates Python packages)
python -m venv venv

# 3. Activate it
# Windows:
venv\Scripts\activate
# Mac/Linux:
source venv/bin/activate

# 4. Install all Python dependencies
pip install -r requirements.txt

# 5. Create your .env file (copy from example)
# Fill in your secrets (see Section 8 for variables)
\end{termbox}

\begin{termbox}
# 6. Run migrations to create database tables
python manage.py makemigrations
python manage.py migrate

# 7. Create a superuser for the Django Admin panel
python manage.py createsuperuser

# 8. Start the development server (runs on port 8000)
python manage.py runserver
\end{termbox}

Backend is now running at: \texttt{http://localhost:8000}\\
Admin panel: \texttt{http://localhost:8000/admin/}

\subsection{Frontend Setup}

\begin{termbox}
# Navigate to frontend directory
cd TaskFlow/frontend

# Install all Node.js dependencies
npm install

# Start the development server (runs on port 5173)
npm run dev
\end{termbox}

Frontend is now running at: \texttt{http://localhost:5173}

\begin{notebox}
In development, the frontend talks to \texttt{http://localhost:8000}. In production, it talks to your live Render URL. This is configured in \texttt{lib/api.js} using environment variables.
\end{notebox}

\subsection{Backend Environment Variables (.env)}

Create a file called \texttt{.env} inside the \texttt{backend/} folder:

\begin{lstlisting}
# SECURITY
SECRET_KEY=your-very-long-random-secret-key-here
DEBUG=True                    # False in production!

# ALLOWED HOSTS
ALLOWED_HOSTS=localhost,127.0.0.1

# DATABASE (for local SQLite, leave DATABASE_URL empty)
DATABASE_URL=                 # leave empty for SQLite
# For PostgreSQL:
# DATABASE_URL=postgresql://user:password@host:5432/dbname

# CORS (frontend URL)
CORS_ALLOWED_ORIGINS=http://localhost:5173

# SESSION DURATION (in seconds)
SESSION_COOKIE_AGE=86400      # 24 hours
\end{lstlisting}

\begin{warnbox}
\textbf{NEVER commit your .env file to Git!} It contains secrets. Make sure \texttt{.env} is listed in your \texttt{.gitignore} file.
\end{warnbox}

\subsection{Frontend Environment Variables (.env)}

Create a file called \texttt{.env} inside the \texttt{frontend/} folder:

\begin{lstlisting}
VITE_API_BASE_URL=http://localhost:8000
\end{lstlisting}

In production, set this to your Render backend URL.


% ════════════════════════════════════════════════════════════════════════════
\section{Deployment Flow}
% ════════════════════════════════════════════════════════════════════════════

\subsection{Infrastructure Overview}

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Service} & \textbf{Platform} & \textbf{Cost} \\
\midrule
Database & Neon.tech & Free tier \\
Backend API & Render & Free tier \\
Frontend & Vercel & Free tier \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Step 1: Set Up the Database (Neon.tech)}

\begin{enumerate}
    \item Go to \url{https://neon.tech} and create a free account
    \item Create a new project (e.g., "taskflow")
    \item Copy the \textbf{Connection String} (it looks like:\\ \texttt{postgresql://user:pass@host.neon.tech/dbname?sslmode=require})
    \item Save this — it is your \texttt{DATABASE\_URL}
\end{enumerate}

\subsection{Step 2: Deploy Backend to Render}

\begin{enumerate}
    \item Push your code to GitHub
    \item Go to \url{https://render.com} and create a free account
    \item Click \textbf{New} → \textbf{Web Service}
    \item Connect your GitHub repository
    \item Configure:
    \begin{itemize}
        \item \textbf{Name:} task-manager-backend
        \item \textbf{Root Directory:} backend
        \item \textbf{Runtime:} Python 3
        \item \textbf{Build Command:} \texttt{pip install -r requirements.txt}
        \item \textbf{Start Command:} \texttt{gunicorn core.wsgi:application}
    \end{itemize}
    \item Add all \textbf{Environment Variables} (from your .env) in Render's dashboard
    \item Set \texttt{DEBUG=False} and \texttt{ALLOWED\_HOSTS} to your Render domain
    \item After deploying, run migrations:
\end{enumerate}

\begin{termbox}
# In Render dashboard, open Shell tab and run:
python manage.py migrate
python manage.py collectstatic --no-input  # serve static files
\end{termbox}

\begin{notebox}
Render automatically restarts your app on every \texttt{git push} to your connected branch.
\end{notebox}

\subsection{Step 3: Deploy Frontend to Vercel}

\begin{enumerate}
    \item Go to \url{https://vercel.com} and create a free account
    \item Click \textbf{New Project} → import your GitHub repo
    \item Configure:
    \begin{itemize}
        \item \textbf{Root Directory:} frontend
        \item \textbf{Framework Preset:} Vite
        \item \textbf{Build Command:} \texttt{npm run build}
        \item \textbf{Output Directory:} dist
    \end{itemize}
    \item Add Environment Variable: \texttt{VITE\_API\_BASE\_URL=https://your-render-url.onrender.com}
    \item Click Deploy
\end{enumerate}

\subsubsection{React Router Fix (vercel.json)}
Without this file, refreshing any page other than \texttt{/} gives a 404 error. Create \texttt{frontend/vercel.json}:

\begin{lstlisting}[language=bash]
{
  "rewrites": [
    { "source": "/(.*)", "destination": "/index.html" }
  ]
}
\end{lstlisting}

This tells Vercel to always serve \texttt{index.html} and let React Router handle the routing.

\subsection{Step 4: Configure Cross-Domain Settings}

Since frontend (Vercel) and backend (Render) are on different domains, the following settings are critical:

\subsubsection{Render Environment Variables}
\begin{lstlisting}
SECRET_KEY=<your-secret-key>
DEBUG=False
DATABASE_URL=postgresql://...   (from Neon.tech)
ALLOWED_HOSTS=your-backend.onrender.com
CORS_ALLOWED_ORIGINS=https://your-frontend.vercel.app
SESSION_COOKIE_AGE=86400
\end{lstlisting}

\subsubsection{Django settings.py Production Behavior}
When \texttt{DEBUG=False}, these settings activate:
\begin{itemize}
    \item \texttt{SESSION\_COOKIE\_SAMESITE = 'None'} — allows cross-domain session cookies
    \item \texttt{SESSION\_COOKIE\_SECURE = True} — HTTPS only
    \item \texttt{CSRF\_COOKIE\_SAMESITE = 'None'} — allows cross-domain CSRF cookies
    \item \texttt{CSRF\_COOKIE\_SECURE = True} — HTTPS only
\end{itemize}

\subsection{Deployment Troubleshooting}

\begin{center}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Problem} & \textbf{Solution} \\
\midrule
403 CSRF Failed & Ensure CORS\_ALLOWED\_ORIGINS includes the frontend URL. Check CSRF\_TRUSTED\_ORIGINS too. \\
403 Not Authenticated & Session cookie not being sent. Check \texttt{withCredentials: true} in Axios and \texttt{CORS\_ALLOW\_CREDENTIALS=True} in Django. \\
SECRET\_KEY error & Set the SECRET\_KEY environment variable in Render dashboard. \\
500 Server Error & Check Render logs. Common causes: missing migrations, missing env vars, import errors. \\
404 on page refresh & Add vercel.json rewrites as shown above. \\
Static files not found & Run \texttt{collectstatic} and ensure Whitenoise is installed. \\
\bottomrule
\end{tabularx}
\end{center}


% ════════════════════════════════════════════════════════════════════════════
\section{Security Considerations}
% ════════════════════════════════════════════════════════════════════════════

\begin{center}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Concern} & \textbf{How We Handle It} \\
\midrule
Password storage & Django hashes with PBKDF2-SHA256. Plain passwords never stored. \\
Session hijacking & \texttt{SESSION\_COOKIE\_HTTPONLY=True} prevents JavaScript from reading the cookie. \\
CSRF attacks & Django's CSRF middleware + manual CSRF token in request headers. \\
SQL Injection & Django ORM parameterizes all queries automatically. \\
Object-level access & Custom \texttt{IsTeamOwner} permission class on team endpoints. \\
Cross-origin access & Only specific origins in \texttt{CORS\_ALLOWED\_ORIGINS}; credentials required. \\
Secret key exposure & SECRET\_KEY loaded from environment variable, raises ValueError if missing. \\
Debug mode in production & DEBUG defaults to \texttt{False} if env variable not set. \\
\bottomrule
\end{tabularx}
\end{center}


% ════════════════════════════════════════════════════════════════════════════
\section{Key Flows — End to End}
% ════════════════════════════════════════════════════════════════════════════

\subsection{User Login Flow}
\begin{enumerate}
    \item User types email + password in \texttt{LoginPage.jsx}
    \item React Hook Form validates locally (non-empty fields)
    \item Axios sends \texttt{POST /api/auth/login/} with \texttt{credentials: 'include'}
    \item Django \texttt{LoginView} calls \texttt{authenticate(email, password)}
    \item If valid: \texttt{login(request, user)} creates session in DB, sets cookie in response
    \item Response includes \texttt{csrf\_token} and \texttt{user} data
    \item Frontend stores user in Zustand, stores csrf\_token in module variable
    \item React Router navigates to \texttt{/} (Dashboard)
\end{enumerate}

\subsection{Create Task Flow}
\begin{enumerate}
    \item User clicks "New Task" → \texttt{TaskModal} opens
    \item User selects a team → frontend fetches \texttt{/api/teams/:id/members/} (staleTime:0 = always fresh)
    \item User fills form and selects assignees
    \item On submit: Axios sends \texttt{POST /api/tasks/} with \texttt{assigned\_to\_ids: [2, 3]}
    \item Django \texttt{TaskViewSet.create()} validates all assignees are team members
    \item Task is saved, a notification is sent to each assignee
    \item React Query invalidates \texttt{['tasks']} cache → TasksPage refreshes automatically
\end{enumerate}

\subsection{Add Team Member Flow}
\begin{enumerate}
    \item Owner clicks "Add Member" in a team card
    \item Types email in \texttt{MemberModal}
    \item Frontend sends \texttt{POST /api/teams/:id/add\_member/} with \texttt{\{email: "...\}}
    \item Django looks up user by email, creates a \texttt{TeamMembership} record
    \item Frontend invalidates \textbf{both} \texttt{['teams']} AND \texttt{['team-members', teamId]}
    \item Next time TaskModal opens for this team, the new member appears immediately — no refresh needed
\end{enumerate}


% ════════════════════════════════════════════════════════════════════════════
\section{Quick Reference}
% ════════════════════════════════════════════════════════════════════════════

\subsection{Common Django Commands}
\begin{termbox}
# Run the dev server
python manage.py runserver

# Create migrations after changing models.py
python manage.py makemigrations

# Apply migrations to database
python manage.py migrate

# Create admin user
python manage.py createsuperuser

# Open Django shell (interactive Python with DB access)
python manage.py shell

# Check for issues
python manage.py check
\end{termbox}

\subsection{Common Frontend Commands}
\begin{termbox}
# Start dev server
npm run dev

# Build for production
npm run build

# Preview production build locally
npm run preview

# Install a new package
npm install package-name
\end{termbox}

\subsection{Git Workflow}
\begin{termbox}
# Stage all changes
git add .

# Commit with a message
git commit -m "feat: add notification system"

# Push to GitHub (triggers Render and Vercel auto-deploy)
git push
\end{termbox}

\subsection{Environment Variable Summary}
\begin{center}
\begin{tabularx}{\textwidth}{llX}
\toprule
\textbf{Variable} & \textbf{Location} & \textbf{Description} \\
\midrule
SECRET\_KEY & backend/.env & Django secret (must be random, long) \\
DEBUG & backend/.env & True in dev, False in prod \\
ALLOWED\_HOSTS & backend/.env & Comma-separated list of allowed domains \\
DATABASE\_URL & backend/.env & Full PostgreSQL connection string \\
CORS\_ALLOWED\_ORIGINS & backend/.env & Frontend URL(s), comma-separated \\
SESSION\_COOKIE\_AGE & backend/.env & Session lifetime in seconds (86400 = 24h) \\
VITE\_API\_BASE\_URL & frontend/.env & Backend API URL \\
\bottomrule
\end{tabularx}
\end{center}


% ════════════════════════════════════════════════════════════════════════════
\section{Glossary}
% ════════════════════════════════════════════════════════════════════════════

\begin{description}
    \item[API] Application Programming Interface — a way for two programs to talk to each other using HTTP requests and JSON responses.
    \item[REST] Representational State Transfer — a design pattern for APIs using standard HTTP methods (GET, POST, PATCH, DELETE).
    \item[JSON] JavaScript Object Notation — a text format for data that both Python and JavaScript can read/write. Looks like: \texttt{\{"key": "value"\}}.
    \item[Model (Django)] A Python class that maps to a database table. Each instance is a row.
    \item[Migration] A file that describes database changes. Run with \texttt{migrate} to apply.
    \item[Serializer] Converts Python objects to JSON (serialization) and validates JSON back to Python (deserialization).
    \item[ViewSet] A Django class that handles multiple API actions (list, create, retrieve, update, delete) in one place.
    \item[Session] A server-side record of a user being logged in. Identified by a cookie sent to the browser.
    \item[CSRF Token] A random secret that proves a request came from your own frontend, not a malicious site.
    \item[CORS] Cross-Origin Resource Sharing — browser security that restricts which external domains can make requests to an API.
    \item[ForeignKey] A field linking two models — one-to-many (one team has many tasks).
    \item[ManyToManyField] A field linking two models — many-to-many (many tasks can have many assignees).
    \item[React Query] Library that manages fetching, caching, and synchronizing server data in React.
    \item[Zustand] A minimal global state manager for React.
    \item[Virtual Environment] An isolated Python installation so packages don't conflict between projects.
    \item[Gunicorn] A production-grade Python web server (replaces Django's built-in dev server).
    \item[Whitenoise] Django middleware that serves static files without a separate file server.
    \item[Vite] A fast build tool and dev server for JavaScript/React projects.
\end{description}


\vfill
\noindent\rule{\linewidth}{0.5pt}\\[0.5em]
{\small\color{textgray} TaskFlow Technical Documentation · v1.0 · March 2026 · Built with Django 5.0 + React 18}

\end{document}
